---
title: "Intro to Vectors in R"
format: live-html
engine: knitr
toc: true
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

## Welcome

A **vector** is R's basic container for a sequence of values of the same
type. This short tutorial covers creating vectors, indexing them, and a
couple of summary functions.

## Creating vectors

The `c()` function ("combine") builds a vector out of individual values:

```{webr}
scores <- c(88, 92, 79, 95, 61)
scores
```

You can pull out one element with `[ ]` and a position (R counts from 1,
not 0):

```{webr}
scores[2]
```

## Exercise: total and average

```{webr}
#| exercise: ex_total
#| persist: true
scores <- c(88, 92, 79, 95, 61)

# Compute the sum of `scores`.
______
```

::: { .hint exercise="ex_total" }
The `sum()` function adds up every element of a vector: `sum(scores)`.
:::

::: { .solution exercise="ex_total" }
```r
sum(scores)
```
:::

```{webr}
#| exercise: ex_average
#| persist: true
scores <- c(88, 92, 79, 95, 61)

# Compute the average (mean) of `scores`.
______
```

::: { .hint exercise="ex_average" }
The `mean()` function averages every element of a vector: `mean(scores)`.
:::

::: { .solution exercise="ex_average" }
```r
mean(scores)
```
:::

## Quick check

```{r}
#| label: quick-check-1
#| echo: false
learnr2::question(
  "What does `scores[2]` return, for `scores <- c(88, 92, 79, 95, 61)`?",
  learnr2::answer("92", correct = TRUE),
  learnr2::answer("88"),
  learnr2::answer("79"),
  learnr2::answer("An error --- R vectors start counting at 0"),
  allow_retry = TRUE
)
```

## Your Answers

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "intro-vectors")
```
