---
title: "Hello learnr2"
subtitle: "A guided tour of learnr2 features"
format: live-html
engine: knitr
toc: true
webr:
  packages:
    - ggplot2
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

{{< include _extensions/r-wasm/live/_gradethis.qmd >}}

## What is this?

**learnr2** lets you write interactive R tutorials that run entirely in the
reader's web browser using [Quarto](https://quarto.org) and
[WebR](https://docs.r-wasm.org/webr/). There is no Shiny app and no R server:
the code cells below execute right here on this page.

This tutorial is itself a learnr2 tutorial. Read it, edit the cells, and run
them to see each feature work.

## 1. Runnable code cells

Every code cell is a live editor. Press **Run Code** (or `Ctrl`/`Cmd` +
`Enter`) to execute it. Edit the code and run it again --- it is yours to experiment with.

```{webr}
#| label: runnable-code-cells-1
x <- c(4, 8, 15, 16, 23, 42)
mean(x)
```

Output works the same as in R, including plots:

```{webr}
#| label: runnable-code-cells-2
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x)
```

## 2. Non-editable cells

Sometimes you want to *show* code that runs but should not be edited. Use the
`edit: false` option.

```{webr}
#| label: non-editable-cells-1
#| edit: false
sample(1:100, 5)
```

## 3. Exercises

An **exercise** is a code cell the reader is expected to complete. Mark a cell
with the `exercise` option and give it a label. The cell starts empty (or
with a leading comment, as below) for the reader to fill in.

```{webr}
#| label: exercises-1
#| exercise: ex_sum
#| persist: true
# Compute the sum of the integers from 1 to 100.
```

Add `persist: true` (as shown above) to save the reader's edits in the
browser's `localStorage`, keyed to this page and exercise. Refresh the page
after editing the cell above --- your changes are still there.

### Hints

Add a `.hint` block tied to the exercise label. Readers can reveal it when
stuck.

::: { .hint exercise="ex_sum" }
The `sum()` function adds numbers, and `1:100` is the sequence 1, 2, ..., 100.
:::

### Solutions

A `.solution` block reveals a model answer.

::: { .solution exercise="ex_sum" }
```r
sum(1:100)
```
:::

## 4. Setup cells

Exercises often need objects to exist before the reader starts. A cell marked
`setup: true` for a given exercise label runs invisibly beforehand. Here we
pre-create a vector the next exercise uses.

```{webr}
#| label: setup-cells-1
#| setup: true
#| exercise: ex_mean
scores <- c(72, 88, 91, 64, 100)
```

```{webr}
#| label: setup-cells-2
#| exercise: ex_mean
#| persist: true
# `scores` already exists. Compute its mean.
```

::: { .solution exercise="ex_mean" }
```r
mean(scores)
```
:::

## 5. Automatic grading

learnr2 reuses 'quarto-live''s grading, which is powered by the
[gradethis](https://pkgs.rstudio.com/gradethis/) package. Add a `check` cell
for an exercise; it runs when the reader presses **Submit** and compares their
work against the solution.

A graded exercise needs its solution as a `solution: true` **code cell**
(not the `.solution` fenced div used in section 3) --- only the cell form
emits the machine-readable copy of the code that `gradethis::grade_this_code()`
compares against. The cell still renders a "Show solution" button for the
reader, exactly like the div does.

```{webr}
#| label: automatic-grading-1
#| exercise: ex_grade
#| persist: true
# Multiply 6 by 7.
```

```{webr}
#| label: automatic-grading-2
#| exercise: ex_grade
#| solution: true
6 * 7
```

```{webr}
#| label: automatic-grading-3
#| exercise: ex_grade
#| check: true
gradethis::grade_this_code()
```

Try submitting a wrong answer (e.g. `6 + 7`), then the right one, to see the
feedback change.

## 6. Quiz questions

Not every check needs to run code. `learnr2::question()` adds a learnr-style
quiz question --- graded entirely in the reader's browser with plain
JavaScript, no Shiny or R server involved. Questions render wherever the
R code below is evaluated (a normal `{r}` chunk, since this runs once while
you *author* the tutorial, not each time a reader loads the page).

Single choice, with a nudge to try again:

```{r}
#| label: quiz-questions-1
learnr2::question(
  "Which function computes the arithmetic mean in base R?",
  learnr2::answer("mean()", correct = TRUE),
  learnr2::answer("average()"),
  learnr2::answer("avg()"),
  allow_retry = TRUE
)
```

Multiple choice is inferred automatically from how many answers are marked
`correct`:

```{r}
#| label: quiz-questions-2
learnr2::question(
  "Which of these are valid ways to create a vector in R?",
  learnr2::answer("c(1, 2, 3)", correct = TRUE),
  learnr2::answer("1:3", correct = TRUE),
  learnr2::answer("vector(1, 2, 3)"),
  random_answer_order = TRUE
)
```

Free-text questions compare the reader's typed answer against one or more
acceptable responses:

```{r}
#| label: quiz-questions-3
learnr2::question(
  "What R package powers grading for learnr2 exercises?",
  learnr2::answer("gradethis", correct = TRUE),
  type = "text",
  allow_retry = TRUE
)
```

Some free-response questions have a definite right answer that's too long or
variable in phrasing to grade with an exact match. `type = "reflection"`
handles that case: it shows the reader a model answer after they submit
their own response, without grading it --- and locks their response so it
can't be changed afterward:

```{r}
#| label: quiz-questions-4
learnr2::question(
  "Explain why the code cells on this page don't need a running R server.",
  learnr2::answer(
    "They execute in the reader's own browser via WebR (R compiled to
    WebAssembly), not on a remote server --- so the tutorial keeps working
    even after wherever it was originally hosted goes offline.",
    correct = TRUE
  ),
  type = "reflection"
)
```

`type = "reflection_editable"` works the same way, except the reader can
keep revising their own response after the model answer is revealed:

```{r}
#| label: quiz-questions-5
learnr2::question(
  "Explain the difference between `type = \"text\"` and `type =
  \"reflection\"`.",
  learnr2::answer(
    "`\"text\"` grades an exact match against one or more accepted answers;
    `\"reflection\"` never grades anything, it just shows a model answer for
    comparison.",
    correct = TRUE
  ),
  type = "reflection_editable"
)
```

Submitted answers are saved automatically (again via `localStorage`, keyed
by page URL and question) and restored on the reader's next visit. The
default key is the question's `{r}` chunk label, so rewording a question
is safe --- but renaming its chunk resets that question's saved answers.
Pass `id = "some-stable-slug"` to `question()` to pin the key regardless
of the chunk label.

Group related questions with `learnr2::quiz()`:

```{r}
#| label: quiz-questions-6
learnr2::quiz(
  caption = "Quick check",
  learnr2::question(
    "2 + 2 equals?",
    learnr2::answer("4", correct = TRUE),
    learnr2::answer("22")
  ),
  learnr2::question(
    "Is R free and open source?",
    learnr2::answer("Yes", correct = TRUE),
    learnr2::answer("No")
  )
)
```

## 7. Copy/paste an image

For exercises where the reader's work is a plot or other image, add
`allow_image = TRUE` to a `"reflection"`/`"reflection_editable"` question
and let them paste a screenshot from their clipboard *as* their response
--- there is no file-upload button, just press Ctrl+V (or Cmd+V) with the
response box focused (or click the box below it and paste there instead).
Once an image is pasted it is the whole answer: the text box disappears
and nothing typed is saved alongside it. "Remove image" brings the text
box back.

Try it: edit the exercise below (or leave it as-is), run it, then take a
screenshot of the plot it produces --- e.g. `Cmd+Shift+4` on a Mac,
`Win+Shift+S` on Windows, or your Chromebook's screenshot tool (`Ctrl` +
`Show windows`) --- and paste it into the question underneath.

```{webr}
#| label: copy-paste-an-image-1
#| exercise: ex_image_plot
#| persist: true
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(color = "steelblue", size = 3)
```

```{r}
#| label: copy-paste-an-image-2
learnr2::question(
  "Paste a screenshot of the plot you made in the exercise above.",
  learnr2::answer(
    "A scatterplot of `wt` vs. `mpg`, colored steel blue.",
    correct = TRUE
  ),
  type = "reflection",
  allow_image = TRUE
)
```

Whatever your OS's native clipboard format actually is behind the scenes
(bitmap, TIFF, etc. --- it genuinely varies by platform, so a screenshot
isn't guaranteed to already be a PNG), `allow_image` accepts it directly
--- PNG, JPEG, GIF, WebP, or BMP --- and re-encodes it as PNG itself before
storing it, so what's actually saved is always PNG no matter which of
those your platform handed over (see the `question()` help page for the
exact size limit).

## 8. Student info and submission

Two more pieces for wrapping a tutorial into something a student can turn
in, no server required. In a real tutorial, `student_info()` normally goes
at the very top and `download_answers_button()` at the very bottom; they're
shown together here just for the demo.

`student_info()` collects identifying information (name, email, an
optional ID by default). It's deliberately *not* a quiz question --- there
is nothing to grade and no model answer, just a plain form with its own
Submit button. It also auto-saves as the reader types, so nothing is lost
even if they never click Submit:

```{r}
#| label: student-info-and-submission-1
learnr2::student_info()
```

`download_answers_button()` gathers every `question()` and `student_info()`
answer already saved on the page into one readable JSON file and downloads
it --- so a reader can save their work and turn it in (attach to an email,
upload to an LMS), all without a server to submit to. The download also
carries browser info, a per-device id, and a `time` field: the moment the
reader clicked "Download", lightly obfuscated so the raw timestamp isn't
readable or hand-editable in the file. Recover it with
`learnr2::submission_time(path)`. Nothing else is hashed or signed --- with
no server involved there is no way to *prevent* a reader editing their
answers; this just keeps the submission time honest at a glance:

```{r}
#| label: student-info-and-submission-2
learnr2::download_answers_button(filename_prefix = "hello-learnr2")
```

## 9. Layout options

Nothing in this section is a widget. It describes the page-level choices an
author has, all made in the YAML header or with one function call, and what
each one does for the reader.

### Table of contents

Every tutorial scaffolded by `create_tutorial()` has `toc: true` in its
YAML header, which gives the page Quarto's table-of-contents sidebar: the
list of sections on the right of this page. Set `toc: false` to drop it.

The sidebar is a map, not a shortcut. Sections reveal one at a time behind
**Continue** buttons whether or not there is a TOC; the TOC adds a view of
the whole tutorial's shape and of how far the reader has come. On a
phone-width screen Quarto hides the sidebar on its own, so a `toc: true`
tutorial behaves like a `toc: false` one there, with no author action.

### Skipping ahead

By default the TOC cannot be used to read ahead. Look at the sidebar now:
the entries below this section are dimmed, a click on one does nothing, and
each lights up only when its section is unlocked. A reader can see what is
coming and where they are, but cannot open the whole tutorial at once.

If you want readers to roam freely --- reference material, or a tour like
this one --- call `learnr2::tutorial_options()` once, in a chunk with
`#| echo: false`, anywhere in the document:

```r
learnr2::tutorial_options(allow_skip = TRUE)
```

Then clicking any TOC entry unlocks every section up to that one and jumps
there, like a classic learnr tutorial with `allow_skip: yes`. This page
leaves the default in place so you can see it in action. Without a TOC the
setting has no effect, since there is nothing to click.

### Pauses without a heading

Not every Continue stop needs a title. A line holding only `###`, with
nothing after it, is a *pause*: the content below it stays hidden until
the reader clicks a button that says just **Continue**, and no heading is
shown. The tutorial.helpers house style puts two of these inside every
exercise, one before the author's own answer and one before the knowledge
drop, so the reader compares before they read on. Here is one now.

###

You just clicked through a pause. In the `.qmd` it is three hash marks on
a line by themselves; in the sidebar it has no entry.

### Start Over

Every answer, pasted image, and exercise edit on a page is saved in the
reader's browser, so reloading never loses anything. The way to wipe it all
is the **Start Over** button, which learnr2 adds to every tutorial
automatically --- nothing to write in the `.qmd`. It asks for
confirmation, then clears this page's saved state (but not the per-device
id that `download_answers_button()` records) and reloads.

Where it appears follows the layout. With a visible sidebar it sits at the
bottom of the table of contents, as it does on this page. Without one ---
`toc: false`, or a screen too narrow for the sidebar --- it moves to the
top of the tutorial, directly under the title, where it stays visible no
matter how far the reader has progressed. Either way the reader always has
exactly one.

## Where to go next

- Use `learnr2::create_tutorial("my-tutorial", dir = ".")` to scaffold
  your own.
- Edit the `.qmd`, then run `learnr2::check_tutorial()` on its directory
  to catch the common authoring mistakes, and
  `learnr2::render_tutorials()` (or plain Quarto) to build it.
- Preview the result with `learnr2::run_tutorial()`, which renders, serves
  and opens it in your browser.
