---
title: "Hello, learnr2"
subtitle: "A guided tour of learnr2 features"
format: live-html
engine: knitr
toc: true
webr:
  packages:
    - ggplot2
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

{{< include _extensions/r-wasm/live/_gradethis.qmd >}}

## What is this?

**learnr2** lets you write interactive R tutorials that run entirely in the
reader's web browser using [Quarto](https://quarto.org) and
[WebR](https://docs.r-wasm.org/webr/). There is no Shiny app and no R server:
the code cells below execute right here on this page.

This tutorial is itself a learnr2 tutorial. Read it, edit the cells, and run
them to see each feature work.

## 1. Runnable code cells

Every code cell is a live editor. Press **Run** (or `Ctrl`/`Cmd` + `Enter`) to
execute it. Edit the code and run it again --- it is yours to experiment with.

```{webr}
#| label: runnable-code-cells-1
x <- c(4, 8, 15, 16, 23, 42)
mean(x)
```

Output works the same as in R, including plots:

```{webr}
#| label: runnable-code-cells-2
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x)
```

## 2. Non-editable cells

Sometimes you want to *show* code that runs but should not be edited. Use the
`edit: false` option.

```{webr}
#| label: non-editable-cells-1
#| edit: false
sample(1:100, 5)
```

## 3. Exercises

An **exercise** is a code cell the reader is expected to complete. Mark a cell
with the `exercise` option and give it a label. The cell starts empty (or
with a leading comment, as below) for the reader to fill in.

```{webr}
#| label: exercises-1
#| exercise: ex_sum
#| persist: true
# Compute the sum of the integers from 1 to 100.
```

Add `persist: true` (as shown above) to save the reader's edits in the
browser's `localStorage`, keyed to this page and exercise. Refresh the page
after editing the cell above --- your changes are still there.

### Hints

Add a `.hint` block tied to the exercise label. Readers can reveal it when
stuck.

::: { .hint exercise="ex_sum" }
The `sum()` function adds numbers, and `1:100` is the sequence 1, 2, ..., 100.
:::

### Solutions

A `.solution` block reveals a model answer.

::: { .solution exercise="ex_sum" }
```r
sum(1:100)
```
:::

## 4. Setup cells

Exercises often need objects to exist before the reader starts. A cell marked
`setup: true` for a given exercise label runs invisibly beforehand. Here we
pre-create a vector the next exercise uses.

```{webr}
#| label: setup-cells-1
#| setup: true
#| exercise: ex_mean
scores <- c(72, 88, 91, 64, 100)
```

```{webr}
#| label: setup-cells-2
#| exercise: ex_mean
#| persist: true
# `scores` already exists. Compute its mean.
```

::: { .solution exercise="ex_mean" }
```r
mean(scores)
```
:::

## 5. Automatic grading

learnr2 reuses 'quarto-live''s grading, which is powered by the
[gradethis](https://pkgs.rstudio.com/gradethis/) package. Add a `check` cell
for an exercise; it runs when the reader presses **Submit** and compares their
work against the solution.

A graded exercise needs its solution as a `solution: true` **code cell**
(not the `.solution` fenced div used in section 3) --- only the cell form
emits the machine-readable copy of the code that `gradethis::grade_this_code()`
compares against. The cell still renders a "Show solution" button for the
reader, exactly like the div does.

```{webr}
#| label: automatic-grading-1
#| exercise: ex_grade
#| persist: true
# Multiply 6 by 7.
```

```{webr}
#| label: automatic-grading-2
#| exercise: ex_grade
#| solution: true
6 * 7
```

```{webr}
#| label: automatic-grading-3
#| exercise: ex_grade
#| check: true
gradethis::grade_this_code()
```

Try submitting a wrong answer (e.g. `6 + 7`), then the right one, to see the
feedback change.

## 6. Quiz questions

Not every check needs to run code. `learnr2::question()` adds a learnr-style
quiz question --- graded entirely in the reader's browser with plain
JavaScript, no Shiny or R server involved. Questions render wherever the
R code below is evaluated (a normal `{r}` chunk, since this runs once while
you *author* the tutorial, not each time a reader loads the page).

Single choice, with a nudge to try again:

```{r}
#| label: quiz-questions-1
learnr2::question(
  "Which function computes the arithmetic mean in base R?",
  learnr2::answer("mean()", correct = TRUE),
  learnr2::answer("average()"),
  learnr2::answer("avg()"),
  allow_retry = TRUE
)
```

Multiple choice is inferred automatically from how many answers are marked
`correct`:

```{r}
#| label: quiz-questions-2
learnr2::question(
  "Which of these are valid ways to create a vector in R?",
  learnr2::answer("c(1, 2, 3)", correct = TRUE),
  learnr2::answer("1:3", correct = TRUE),
  learnr2::answer("vector(1, 2, 3)"),
  random_answer_order = TRUE
)
```

Free-text questions compare the reader's typed answer against one or more
acceptable responses:

```{r}
#| label: quiz-questions-3
learnr2::question(
  "What R package powers grading for learnr2 exercises?",
  learnr2::answer("gradethis", correct = TRUE),
  type = "text",
  allow_retry = TRUE
)
```

Some free-response questions have a definite right answer that's too long or
variable in phrasing to grade with an exact match. `type = "reflection"`
handles that case: it shows the reader a model answer after they submit
their own response, without grading it --- and locks their response so it
can't be changed afterward:

```{r}
#| label: quiz-questions-4
learnr2::question(
  "Explain why the code cells on this page don't need a running R server.",
  learnr2::answer(
    "They execute in the reader's own browser via WebR (R compiled to
    WebAssembly), not on a remote server --- so the tutorial keeps working
    even after wherever it was originally hosted goes offline.",
    correct = TRUE
  ),
  type = "reflection"
)
```

`type = "reflection_editable"` works the same way, except the reader can
keep revising their own response after the model answer is revealed:

```{r}
#| label: quiz-questions-5
learnr2::question(
  "Explain the difference between `type = \"text\"` and `type =
  \"reflection\"`.",
  learnr2::answer(
    "`\"text\"` grades an exact match against one or more accepted answers;
    `\"reflection\"` never grades anything, it just shows a model answer for
    comparison.",
    correct = TRUE
  ),
  type = "reflection_editable"
)
```

Submitted answers are saved automatically (again via `localStorage`, keyed
by page URL and question) and restored on the reader's next visit. The
default key is the question's `{r}` chunk label, so rewording a question
is safe --- but renaming its chunk resets that question's saved answers.
Pass `id = "some-stable-slug"` to `question()` to pin the key regardless
of the chunk label.

Group related questions with `learnr2::quiz()`:

```{r}
#| label: quiz-questions-6
learnr2::quiz(
  caption = "Quick check",
  learnr2::question(
    "2 + 2 equals?",
    learnr2::answer("4", correct = TRUE),
    learnr2::answer("22")
  ),
  learnr2::question(
    "Is R free and open source?",
    learnr2::answer("Yes", correct = TRUE),
    learnr2::answer("No")
  )
)
```

## 7. Copy/paste an image

For exercises where the reader's work is a plot or other image, add
`allow_image = TRUE` to a `"reflection"`/`"reflection_editable"` question
and let them paste a screenshot from their clipboard alongside their
written response --- there is no file-upload button, just press Ctrl+V (or
Cmd+V) with the response box focused (or click the box below it and paste
there instead).

Try it: edit the exercise below (or leave it as-is), run it, then take a
screenshot of the plot it produces --- e.g. `Cmd+Shift+4` on a Mac,
`Win+Shift+S` on Windows, or your Chromebook's screenshot tool (`Ctrl` +
`Show windows`) --- and paste it into the question underneath.

```{webr}
#| label: copy-paste-an-image-1
#| exercise: ex_image_plot
#| persist: true
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(color = "steelblue", size = 3)
```

```{r}
#| label: copy-paste-an-image-2
learnr2::question(
  "Paste a screenshot of the plot you made in the exercise above.",
  learnr2::answer(
    "A scatterplot of `wt` vs. `mpg`, colored steel blue.",
    correct = TRUE
  ),
  type = "reflection",
  allow_image = TRUE
)
```

Whatever your OS's native clipboard format actually is behind the scenes
(bitmap, TIFF, etc. --- it genuinely varies by platform, so a screenshot
isn't guaranteed to already be a PNG), `allow_image` accepts it directly
--- PNG, JPEG, GIF, WebP, or BMP --- and re-encodes it as PNG itself before
storing it, so what's actually saved is always PNG no matter which of
those your platform handed over (see [question()]'s docs for the exact
size limit).

## 8. Student info and submission

Two more pieces for wrapping a tutorial into something a student can turn
in, no server required. In a real tutorial, `student_info()` normally goes
at the very top and `download_answers_button()` at the very bottom; they're
shown together here just for the demo.

`student_info()` collects identifying information (name, email, an
optional ID by default). It's deliberately *not* a quiz question --- there
is nothing to grade and no model answer, just a plain form with its own
Submit button. It also auto-saves as the reader types, so nothing is lost
even if they never click Submit:

```{r}
#| label: student-info-and-submission-1
learnr2::student_info()
```

`download_answers_button()` gathers every `question()` and `student_info()`
answer already saved on the page into one readable JSON file and downloads
it --- so a reader can save their work and turn it in (attach to an email,
upload to an LMS), all without a server to submit to. The download also
carries browser info, a per-device id, and a `time` field: the moment the
reader clicked "Download", lightly obfuscated so the raw timestamp isn't
readable or hand-editable in the file. Recover it with
`learnr2::submission_time(path)`. Nothing else is hashed or signed --- with
no server involved there is no way to *prevent* a reader editing their
answers; this just keeps the submission time honest at a glance:

```{r}
#| label: student-info-and-submission-2
learnr2::download_answers_button(filename_prefix = "hello-learnr2")
```

## Where to go next

- Use `learnr2::create_tutorial("my-tutorial")` to scaffold your own.
- Edit the `.qmd`, then render with Quarto or `quarto::quarto_render()`.

A learnr-style progressive, section-by-section theme is on the roadmap.
