---
title: "Getting Started"
subtitle: "A five-minute introduction to how these tutorials work"
format: live-html
engine: knitr
toc: true
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

## Introduction

This tutorial teaches you the mechanics of working through a **learnr2**
tutorial --- not R itself. It is meant to be the first tutorial you ever
complete, before any that actually teach R.

There are three steps to every tutorial: tell us who you are, work through
the exercises, and download your answers when you are done. This one is no
different, and should take about five minutes.

A learnr2 tutorial runs entirely in your web browser using
[Quarto](https://quarto.org) and [WebR](https://docs.r-wasm.org/webr/) (R
compiled to WebAssembly). There is no separate R or RStudio installation to
set up, and no server involved --- the code cells on this page run right
here as you work through them.

Start with your name and email below. (Your instructor might also ask you
to fill in the "ID" field so your work can be matched to a grading roster;
if not, just leave it blank.) Fill in the fields and click **Submit** ---
and don't worry about getting it exactly right the first time: your answers
save automatically as you type, and clicking back into a field and changing
it updates what's saved, mistakes included.

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

## Running R Code

Every learnr2 tutorial has cells like the one below, where you write and
run R code right here in your browser, alongside prose questions like the
"Submit Answer" one you just used for your name and email.

### Exercise 1

If you were sitting at a real R console, one of the first things worth
trying is `getwd()`, which asks R "what folder are you currently working
in?" Type `getwd()` into the cell below and click **Run Code**.

```{webr}
#| label: running-r-code-1
#| exercise: exercise_1
#| persist: true
```

::: { .hint exercise="exercise_1" }
getwd()
:::

You should see a path appear below the cell, something like
`[1] "/home/web_user"`. That's not a folder on your own computer --- WebR
runs inside a small virtual file system that lives entirely in your
browser, so don't worry if your path looks different from a friend's. What
matters is that a real command produced a real result, right here, with
nothing to install.

Now answer the question below: paste in both the code you ran and the
result it printed. Get in the habit of copying both the command and its
response together whenever a tutorial (or a colleague) asks what happened
when you ran something --- one without the other is only half the story.

```{r}
#| label: running-r-code-2
#| echo: false
learnr2::question(
  "Paste the code you ran above, and the result it printed.",
  learnr2::answer(
    "Code: getwd(). Result: a path in quotes, e.g. [1] \"/home/web_user\" ---
    the exact path doesn't matter, only that both the command and its
    output are here.",
    correct = TRUE
  ),
  type = "reflection"
)
```

### Exercise 2

Every answer on this page, including the one you just gave, is saved to
your browser as you go --- not only once you finally download your work.
To see this for yourself: reload this page right now. Your name, email,
and the answer above should all still be there, exactly as you left them.

This is also why there's a **Start Over** button at the bottom of the
sidebar rather than relying on a plain reload to clear things: reloading
on its own does not erase anything, on purpose, so you can close a
tutorial partway through and pick it up again later without losing
progress. Only Start Over does that, and only after asking you to confirm.

### Exercise 3

Not every answer is text. Run the code below --- it draws a simple plot ---
then take a screenshot of it (e.g. `Cmd+Shift+4` on a Mac, `Win+Shift+S` on
Windows, or your Chromebook's screenshot tool) and paste it directly into
the question underneath: click into the response box and press Ctrl+V (or
Cmd+V). There's no upload button; pasting is the whole mechanism.

```{webr}
#| label: running-r-code-3
#| exercise: exercise_3
#| persist: true
plot(1:10, 1:10)
```

```{r}
#| label: running-r-code-4
#| echo: false
learnr2::question(
  "Paste a screenshot of the plot you made above.",
  learnr2::answer(
    "A simple scatterplot going from the bottom-left to the top-right corner.",
    correct = TRUE
  ),
  type = "reflection",
  allow_image = TRUE
)
```

## Your Answers

At the end of every tutorial, you will save your work (and submit it, at
your instructor's discretion). First, answer the question below about how
many minutes you spent on the tutorial. Then click the button below that to
gather every answer on this page into a single file and download it.

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "getting-started")
```

The downloaded file's name will start with `getting-started`, followed by a
timestamp, and will be saved wherever your browser saves downloads.
Unless your instructor tells you otherwise, send them this file as-is (for
example, attached to an email or uploaded to your course's learning
management system).

## Summary

That's the whole workflow: tell us who you are, work through the
exercises --- running code, answering prose and quiz questions, all of it
saved automatically as you go --- and download your answers when you're
done. Every other learnr2 tutorial you take follows the same three steps;
only the R content in the middle changes. Good luck!
