---
title: "Getting Started with learnr2 Tutorials"
subtitle: "How exercises, quizzes, and downloads work in this format"
format: live-html
engine: knitr
toc: true
---

{{< include _extensions/r-wasm/live/_knitr.qmd >}}

## Introduction

This tutorial is an introduction to the general use of interactive R
tutorials built with **learnr2**. It is meant for beginning students who
have never completed one before.

A learnr2 tutorial runs entirely in your web browser using
[Quarto](https://quarto.org) and [WebR](https://docs.r-wasm.org/webr/) (R
compiled to WebAssembly). There is no separate R or RStudio installation to
set up, and no server involved --- the code cells on this page run right
here as you work through them. When you are done, you will download a file
containing your answers and then submit it to your instructor (or just keep
a copy for yourself).

Most tutorials begin by asking for your name and email. (Your instructor
might also ask you to fill in the "ID" field so your work can be matched to
a grading roster; if not, just leave it blank.) Fill in the fields below and
click **Submit** --- your answers also save automatically as you type, so
nothing is lost even if you forget to click it.

## Student Information

```{r}
#| label: student-information-1
#| echo: false
learnr2::student_info()
```

## Exercises

In each tutorial, there will be a set of exercises for you to complete. For
non-coding questions, you write your answer and then click the "Submit
Answer" button, as you did above. For coding exercises, you write R code in
a runnable cell and click "Run Code" to see the result.

If you don't click "Run Code" (or "Submit Answer"), your response isn't
recorded.

### Exercise 1

Let's practice some simple math. Type `4 + 5` into the code cell below and
click **Run Code** to see the result. Clicking **Hint** gives you a hint
about what to type.

```{webr}
#| exercise: exercise_1
#| persist: true
______
```

::: { .hint exercise="exercise_1" }
4 + 5
:::

Once you run the code, the answer `[1] 9` should appear below the cell ---
right here on the page, with no separate console to switch to.

There are two types of answers for tutorial questions: prose and code. With
prose answers, you use the "Submit Answer" button, as you did with your
name and email above. With code answers, you use the "Run Code" button, as
you did for this exercise.

### Exercise 2

Let's use variables to perform the same calculation. Below, we have already
assigned 5 to `x` and 4 to `y`. On the blank line, add the two variables by
typing `x + y`. Click **Run Code** to see your result.

```{webr}
#| exercise: exercise_2
#| persist: true
x <- 5
y <- 4

______
```

::: { .hint exercise="exercise_2" }
x + y
:::

Once you run the code, the answer `[1] 9` should appear below the cell.

### Exercise 3

Now change the values to `x <- 6` and `y <- 10` and add them together.
Click **Run Code**. You should get the answer `[1] 16`.

```{webr}
#| exercise: exercise_3
#| persist: true
______
```

::: { .hint exercise="exercise_3" }
x <- ...
y <- 10
x + ...
:::

Note that the assignment symbol, `<-`, is a less-than sign followed by a
dash. We often pronounce it as "gets." So we might say, when reading this
code, that "x gets 6."

If you used the hint, note how we did not provide the entire answer. This
is our standard approach with hints.

Your progress on every exercise and question in this tutorial is saved
automatically to your browser's storage as you work. Try reloading this
page --- your answers above are still there. It's not a problem to close a
tutorial partway through and come back to it later.

### Exercise 4

Let's run a command that reports the version of R powering the code cells
on this page.

```{webr}
#| exercise: exercise_4
#| persist: true
______
```

::: { .hint exercise="exercise_4" }
version[['version.string']]
:::

You should see something like `[1] "R version 4.4.2 (2024-10-31)"`. Because
this R is compiled to WebAssembly and running right here in your browser,
the exact version may differ from what you'd see on a desktop install of R
--- and that's fine. There's no separate R Console involved; the command
runs and reports its result right here, the same way every other exercise
on this page does.

### Exercise 5

Not every exercise involves code. Sometimes we'll ask you to write a
sentence or two of prose instead, with no strict right answer to match.

```{r}
#| label: exercises-1
#| echo: false
learnr2::question(
  "In your own words, what is the difference between a coding exercise
  (\"Run Code\") and a quiz or prose question (\"Submit Answer\") in this
  tutorial?",
  learnr2::answer(
    "A coding exercise runs R code that you write, in a code cell, right
    here in the browser. A quiz or prose question instead asks you to
    choose or type an answer, which is checked (or simply recorded)
    without running any code.",
    correct = TRUE
  ),
  type = "reflection"
)
```

Most instructors grade tutorials on an honest-effort, pass/fail basis, so
don't stress about getting prose answers "right."

### Exercise 6

Some exercises produce a plot instead of a number or a sentence. Let's make
one, using `cars`, a small dataset built into R (speed vs. stopping
distance for cars in the 1920s --- no package needed). Type `plot(cars)`
into the code cell below and click **Run Code**.

```{webr}
#| exercise: exercise_6
#| persist: true
______
```

::: { .hint exercise="exercise_6" }
plot(cars)
:::

You should see a scatterplot appear below the cell, the same way text
output did in the earlier exercises.

Now paste a screenshot of that plot into the box below --- take a
screenshot with your usual tool (e.g. `Cmd+Shift+4` on a Mac, the Snipping
Tool or `Win+Shift+S` on Windows), then click the box and press
`Ctrl+V`/`Cmd+V` to paste it in, as a PNG. There is no file-upload button
to click; pasting is the whole mechanism.

```{r}
#| label: exercises-2
#| echo: false
learnr2::question(
  "Paste a screenshot of the plot you made above.",
  learnr2::answer(
    "A scatterplot with speed on the x-axis and stopping distance on the
    y-axis, trending upward --- faster cars take longer to stop.",
    correct = TRUE
  ),
  type = "reflection",
  allow_image = TRUE
)
```

## Your Answers

At the end of every tutorial, you will save your work (and submit it, at
your instructor's discretion). First, answer the question below about how
many minutes you spent on the tutorial. Then click the button below that to
gather every answer on this page into a single file and download it.

```{r}
#| label: your-answers-1
#| echo: false
learnr2::question(
  "How many minutes, approximately, did it take you to complete this
  tutorial? For example, an hour and a half would be 90 minutes.",
  type = "reflection_editable",
  validate = "integer"
)
```

```{r}
#| label: your-answers-2
#| echo: false
learnr2::download_answers_button(filename_prefix = "getting-started")
```

The downloaded file's name will start with `getting-started`, followed by a
timestamp, and will be saved wherever your browser saves downloads.
Unless your instructor tells you otherwise, send them this file as-is (for
example, attached to an email or uploaded to your course's learning
management system).

## Summary

You now understand how learnr2 tutorials work: running code in live cells
right in your browser, answering quiz and prose questions, and downloading
your answers when you're done. Good luck with your data science journey!
